---
title: "6306 Case Study 2 - "
author: "Andrew Walch, Christopher Morgan, Luke Pierce, Eduardo Cantu Medellin"
date: "April 7, 2018"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Exploratory Data Analysis


A summary done on the data main data yields that some of the fields may not be needed:
  EmployeeCount: This shows 1 for all of them
  EmployeeNumber: This has to do with a random number
  Over18: All of them are over 18
  StandardHours: All records show 80 hours a week.


```{r EDA}
# Load Library 
library("lattice")

# Load Data
attData <- read.csv("CaseStudy2data.csv")
summary(attData)
```

Also, the data was scanned for 'NA' values. This scanned did not show any 'NA' values.
```{r NA_Scan}
# NA per Field
NACount <- sapply(attData,function(x) sum(is.na(x)))
NACount <- as.data.frame(NACount)
colnames(NACount)=c("CountOfNAs")
NACount


```
### Charts



## Conclusion

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
