---
title: "6306 Case Study 2 - "
author: "Andrew Walch, Christopher Morgan, Luke Pierce, Eduardo Cantu Medellin"
date: "April 7, 2018"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}



knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Data Input and Cleaning


A summary done on the data main data yields that some of the fields may not be needed:
  EmployeeCount: This shows 1 for all of them
  EmployeeNumber: This has to do with a random number
  Over18: All of them are over 18
  StandardHours: All records show 80 hours a week.





```{r EDA}

library(pROC)
library(ROCR)
library(randomForest)
library(ggplot2)
library(reshape2)
library(caret)
# Add libraries
library(dplyr)
library(RColorBrewer)
library(forcats)
library(plyr)
library(knitr)
library(kableExtra)
library(data.table)
#library(lattice)
#library(sm)
# Load Data
df <- read.csv("CaseStudy2data.csv")
dfCat <- df[,names(base::Filter(is.factor,df))]
dfCon <- df[,c("Attrition",names(base::Filter(is.integer,df)))]
head(dfCat)
summary(df)

# Set RMarkdown plots to a default 10 by 6 inches
knitr::opts_chunk$set(fig.width=10, fig.height=6)


```

The data shows an inbalance in the number of people that leave the company vs people that stayed in the job. There are 237 samples with attrition 'Yes' and 1233 with attrition 'No'. The data will be down sampled for the Attrition "Yes" to match the level of Attrition "Yes"

Also, the data was scanned for 'NA' values. This scanned did not show any 'NA' values.


```{r NA_Scan}
# NA per Field
NACount <- sapply(df,function(x) sum(is.na(x)))
NACount <- as.data.frame(NACount)
colnames(NACount)=c("CountOfNAs")
NACount


#remove reduntant fields

#remove reduntant fields
# 9 - Employee Count
# 10 - Employee Number
# 22 - Over18
# 27 - Standard Hours
cleanAttDf <- df[,c(-9,-10,-22,-27)]
colnames(cleanAttDf)

# change variables as factors
cleanAttDf$JobLevel <- as.factor(cleanAttDf$JobLevel)
cleanAttDf$JobInvolvement <- as.factor(cleanAttDf$JobInvolvement)
cleanAttDf$EnvironmentSatisfaction <- as.factor(cleanAttDf$EnvironmentSatisfaction)
cleanAttDf$Education <- as.factor(cleanAttDf$Education)
cleanAttDf$JobSatisfaction <- as.factor(cleanAttDf$JobSatisfaction)
cleanAttDf$RelationshipSatisfaction <- as.factor(cleanAttDf$RelationshipSatisfaction)
cleanAttDf$StockOptionLevel <- as.factor(cleanAttDf$StockOptionLevel)
cleanAttDf$PerformanceRating <- as.factor(cleanAttDf$PerformanceRating)
cleanAttDf$WorkLifeBalance <- as.factor(cleanAttDf$WorkLifeBalance)

##### Branch for categorical histograms
orig_cleanAttDf <- cleanAttDf
orig_dfCat <- orig_cleanAttDf[,names(base::Filter(is.factor,orig_cleanAttDf))]
```

## Exploratory Data Analysis 

### Attrition Levels
#### Total attrition levels are significant at 16%.

```{r Attrition, echo=FALSE}

AttritionTable <- table(cleanAttDf$Attrition)
df <- as.data.frame(prop.table(AttritionTable))
df <- plyr::rename(df, c("Var1"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# ggplot Bar Plot (percentage)
g <- ggplot(cleanAttDf, aes(x = Attrition)) + geom_bar(aes(fill=Attrition, y=100*(..count..)/sum(..count..))) 
AttritionBCP <- g + xlab("Attrition") + ylab("Percent") + ggtitle("Total Attrition") 
AttritionBCP
```
### Attrition by Age
#### Almost half (47%) of the 18-22 age bracket leave the company. There is a significant decline with increasing age.

```{r Age, echo=FALSE}
# Subset Age by 5, categorical vector
# 18-22,23-27,28-32,33-37,38-42,43-47,48-52,53+
cleanAttDf$Ageby5 <- findInterval(cleanAttDf$Age, c(18, 23, 28, 33, 38, 43, 48, 53))
cleanAttDf$Ageby5 <- mapvalues(cleanAttDf$Ageby5, from=c(1, 2, 3, 4, 5, 6, 7, 8), to=c('18-22', '23-27', '28-32', '33-37', '38-42', '43-47','48-52','53+'))
cleanAttDf$Ageby5 <- as.factor(cleanAttDf$Ageby5)

# Generate property summary for variable
AgeTable <- table(cleanAttDf$Ageby5, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(AgeTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "Age", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(AgeTable))
df <- plyr::rename(df, c("Var1"= "Age", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = Age, y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
AgeBP <- g + xlab("Age") + ylab("Percent") + ggtitle("Attrition by Age")
AgeBP

```

### Attrition vs. Business Travel
#### Business Travel seems to have little correlation with attrition.

```{r Business Travel, echo=FALSE}
# Generate property summary for variable
BusinessTravelTable <- table(cleanAttDf$BusinessTravel, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(BusinessTravelTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "BusinessTravel", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(BusinessTravelTable))
df <- plyr::rename(df, c("Var1"= "BusinessTravel", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(BusinessTravel, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
BusinessTravelBP <- g + xlab("Business Travel") + ylab("Percent") + ggtitle("Attrition vs. Business Travel")
BusinessTravelBP

```
### Attrition vs. Daily Compensation Rate
#### Daily compensation rate shows some correlation with attrition.

```{r Daily Rate, echo=FALSE}
## Subset Daily Rate by 200, categorical vector
cleanAttDf$DailyRateby200 <- findInterval(cleanAttDf$DailyRate, c(0,200,400,600,800,1000,1200,1400))
cleanAttDf$DailyRateby200 <- mapvalues(cleanAttDf$DailyRateby200, from=c(1, 2, 3, 4, 5, 6, 7, 8), to=c('0-199' ,'200-399', '400-599', '600-799', '800-999', '1000-1199', '1200-1399','1400+'))
cleanAttDf$DailyRateby200 <- as.factor(cleanAttDf$DailyRateby200)

# Generate property summary for variable
DailyRateby200Table <- table(cleanAttDf$DailyRateby200, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(DailyRateby200Table))
PTdf <- plyr::rename(PTdf, c("Var1"= "DailyRateby200", "Var2"= "Attrition"))
levels(PTdf$DailyRateby200) <- c('0-199' ,'200-399', '400-599', '600-799', '800-999', '1000-1199', '1200-1399','1400+')
df <- prop.table(DailyRateby200Table)

kable(PTdf, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = (DailyRateby200), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
DailyRateby200BP <- g + xlab("Daily Rate ($)") + ylab("Percent") + ggtitle("Attrition vs.Daily Rate")
DailyRateby200BP

```
### Attrition by Department
#### Proportionally, HR shows a higher rate of attrition, but the employee counts are low.

```{r Department, echo=FALSE}
# Generate property summary for variable
DepartmentTable <- table(cleanAttDf$Department, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(DepartmentTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "Department", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(BusinessTravelTable))
df <- plyr::rename(df, c("Var1"= "Department", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(Department, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
DepartmentBP <- g + xlab("Department") + ylab("Percent") + ggtitle("Attrition by Department")
DepartmentBP

```
### Attrition vs. Distance From Home
#### Distance from home does not seem to correlate with attrition.

```{r DistanceFromHome, echo=FALSE}

## Subset Distance by 5, categorical vector
# '1-5', '6-10', '11-15', '16-20', '21-25', '25+' 
cleanAttDf$DistanceFromHomeby5 <- findInterval(cleanAttDf$DistanceFromHome, c(0, 5, 10, 15, 20, 25))
cleanAttDf$DistanceFromHomeby5 <- mapvalues(cleanAttDf$DistanceFromHomeby5, from=c(1, 2, 3, 4, 5, 6), to=c('01-05', '06-10', '11-15', '16-20', '21-25', '25+'))
cleanAttDf$DistanceFromHomeby5 <- as.factor(cleanAttDf$DistanceFromHomeby5)

# Generate property summary for variable
DistanceFromHomeby5Table <- table(cleanAttDf$DistanceFromHomeby5, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(DistanceFromHomeby5Table))
PTdf <- plyr::rename(PTdf, c("Var1"= "DistanceFromHomeby5", "Var2"= "Attrition"))
df <- prop.table(DistanceFromHomeby5Table)

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = DistanceFromHomeby5, y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
DistanceFromHomeby5BP <- g + xlab("Distance From Home(miles)") + ylab("Percent") + ggtitle("Attrition vs. Distance From Home")
DistanceFromHomeby5BP

```
### Attrition vs. Education
#### Education has little influence on attrition.

```{r Education, echo=FALSE}
# Generate property summary for variable
EducationTable <- table(cleanAttDf$Education, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(EducationTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "Education", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(EducationTable))
df <- plyr::rename(df, c("Var1"= "Education", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(Education, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
EducationBP <- g + xlab("Education") + ylab("Percent") + ggtitle("Attrition vs. Education")
EducationBP

```

### Attrition vs. Education Field
#### Education Field also has little influence on attrition.

```{r EducationField, echo=FALSE}
# Generate property summary for variable
EducationFieldTable <- table(cleanAttDf$EducationField, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(EducationFieldTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "EducationField", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(EducationFieldTable))
df <- plyr::rename(df, c("Var1"= "EducationField", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(EducationField, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
BusinessTravelBP <- g + xlab("Education Field") + ylab("Percent") + ggtitle("Attrition vs. Education Field")
BusinessTravelBP

```
### Attrition vs. Environment Satisfaction
#### Low Environment Satisfaction seems to have a small effect on attrition.

```{r Environment Satisfaction, echo=FALSE}
# Generate property summary for variable
EnvironmentSatisfactionTable <- table(cleanAttDf$EnvironmentSatisfaction, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(EnvironmentSatisfactionTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "EnvironmentSatisfaction", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(EnvironmentSatisfactionTable))
df <- plyr::rename(df, c("Var1"= "EnvironmentSatisfaction", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(EnvironmentSatisfaction, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
EnvironmentSatisfactionBP <- g + xlab("Environment Satisfaction") + ylab("Percent") + ggtitle("Attrition vs. Environment Satisfaction")
EnvironmentSatisfactionBP

```

### Attrition by Gender
#### Gender seems to have little influence on attrition.

```{r Gender, echo=FALSE}
# Generate property summary for variable
GenderTable <- table(cleanAttDf$Gender, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(GenderTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "Gender", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(GenderTable))
df <- plyr::rename(df, c("Var1"= "Gender", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(Gender, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
GenderBP <- g + xlab("Gender") + ylab("Percent") + ggtitle("Attrition by Gender")
GenderBP

```

### Attrition vs. Hourly Rate
#### Lower hourly rates seem to have influence on increased attrition.

```{r Hourly Rate, echo=FALSE}

## Subset Hourly Rate by 200, categorical vector
cleanAttDf$HourlyRateby10 <- findInterval(cleanAttDf$HourlyRate, c(30, 40, 50, 60, 70, 80, 90))
cleanAttDf$HourlyRateby10 <- mapvalues(cleanAttDf$HourlyRateby10, from=c(1, 2, 3, 4, 5, 6, 7), to=c('30-39' ,'40-49', '50-59', '60-69', '70-79', '80-89', '90-100'))
cleanAttDf$HourlyRateby10 <- as.factor(cleanAttDf$HourlyRateby10)

# Generate property summary for variable
HourlyRateby10Table <- table(cleanAttDf$HourlyRateby10, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(HourlyRateby10Table))
PTdf <- plyr::rename(PTdf, c("Var1"= "HourlyRateby10", "Var2"= "Attrition"))
levels(PTdf$HourlyRateby10) <- c('30-39' ,'40-49', '50-59', '60-69', '70-79', '80-89', '90-100')
df <- prop.table(HourlyRateby10Table)

kable(PTdf, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = HourlyRateby10, y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
HourlyRateby10BP <- g + xlab("Hourly Rate") + ylab("Percent") + ggtitle("Attrition vs. Hourly Rate")
HourlyRateby10BP

```
### Attrition vs. Job Involvement
#### Low Job Involvement does seem to correlate with increased attrition.

```{r Job Involvement, echo=FALSE}
# Generate property summary for variable
JobInvolvementTable <- table(cleanAttDf$JobInvolvement, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(JobInvolvementTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "JobInvolvement", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(JobInvolvementTable))
df <- plyr::rename(df, c("Var1"= "JobInvolvement", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(JobInvolvement, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
JobInvolvementBP <- g + xlab("Job Involvement") + ylab("Percent") + ggtitle("Attrition vs. Job Involvement")
JobInvolvementBP

```

### Attrition vs. Job Level
####  Low Job Level does seem correlated with increased attrition.

```{r Job Level, echo=FALSE}
# Generate property summary for variable
JobLevelTable <- table(cleanAttDf$JobLevel, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(JobLevelTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "JobLevel", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(JobLevelTable))
df <- plyr::rename(df, c("Var1"= "JobLevel", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(JobLevel, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
JobLevelBP <- g + xlab("Job Level") + ylab("Percent") + ggtitle("Attrition vs. Job Level")
JobLevelBP

```

### Attrition vs. Job Role
#### Lab technicians and Sales Representatives seem somewhat more prone to increased attrition.

```{r Job Role, echo=FALSE}
# Generate property summary for variable
JobRoleTable <- table(cleanAttDf$JobRole, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(JobRoleTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "JobRole", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(JobRoleTable))
df <- plyr::rename(df, c("Var1"= "JobRole", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(JobRole, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
g <- g + xlab("Job Role") + ylab("Percent") + ggtitle("Attrition vs. Job Role")
JobRoleBP <- g + theme(axis.text.x = element_text(angle = 90, hjust = 1))
JobRoleBP

```

### Attrition vs. Job Satisfaction
#### Job Satisfaction does seem correlated with increased attrition.

```{r Job Satisfaction, echo=FALSE}
# Generate property summary for variable
JobSatisfactionTable <- table(cleanAttDf$JobSatisfaction, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(JobSatisfactionTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "JobSatisfaction", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(JobSatisfactionTable))
df <- plyr::rename(df, c("Var1"= "JobSatisfaction", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(JobSatisfaction, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
JobSatisfactionBP <- g + xlab("Job Satisfaction") + ylab("Percent") + ggtitle("Attrition vs. Job Satisfaction")
JobSatisfactionBP

```

### Attrition vs. Marital Status
#### Singles seem somewhat more prone to increased attrition, perhaps due to younger age.

```{r Marital Status, echo=FALSE}
# Generate property summary for variable
MaritalStatusTable <- table(cleanAttDf$MaritalStatus, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(MaritalStatusTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "MaritalStatus", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(MaritalStatusTable))
df <- plyr::rename(df, c("Var1"= "MaritalStatus", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(MaritalStatus, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
MaritalStatusBP <- g + xlab("Marital Status") + ylab("Percent") + ggtitle("Attrition vs. Marital Status")
MaritalStatusBP

```

### Attrition vs. Number Companies Worked
#### Number Companies Worked seems to have little impact on attrition.

```{r NumCompaniesWorked, echo=FALSE}
# Generate property summary for variable
NumCompaniesWorkedTable <- table(cleanAttDf$NumCompaniesWorked, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(NumCompaniesWorkedTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "NumCompaniesWorked", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(NumCompaniesWorkedTable))
df <- plyr::rename(df, c("Var1"= "NumCompaniesWorked", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(NumCompaniesWorked, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
NumCompaniesWorkedBP <- g + xlab("Number Companies Worked") + ylab("Percent") + ggtitle("Attrition vs. Number Companies Worked")
NumCompaniesWorkedBP

```

### Attrition vs. OverTime
#### More overtime seems correlated with increased attrition.

```{r OverTime, echo=FALSE}
# Generate property summary for variable
OverTimeTable <- table(cleanAttDf$OverTime, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(OverTimeTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "OverTime", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(OverTimeTable))
df <- plyr::rename(df, c("Var1"= "OverTime", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(OverTime, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
OverTimeBP <- g + xlab("OverTime") + ylab("Percent") + ggtitle("Attrition vs. OverTime")
OverTimeBP

```

### Attrition vs. Percent Salary Hike
#### Suprisingly, salary hikes seem to have little influence on attrition.

```{r Percent Salary Hike, echo=FALSE}
# Generate property summary for variable
PercentSalaryHikeTable <- table(cleanAttDf$PercentSalaryHike, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(PercentSalaryHikeTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "PercentSalaryHike", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(PercentSalaryHikeTable))
df <- plyr::rename(df, c("Var1"= "PercentSalaryHike", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = (PercentSalaryHike), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
PercentSalaryHikeBP <- g + xlab("Percent Salary Hike") + ylab("Percent") + ggtitle("Attrition vs. Percent Salary Hike")
PercentSalaryHikeBP

```

### Attrition vs. Performance Rating
#### Performance Rating also seems to have little influence on attrition.

```{r Performance Rating, echo=FALSE}
# Generate property summary for variable
PerformanceRatingTable <- table(cleanAttDf$PerformanceRating, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(PerformanceRatingTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "PerformanceRating", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(PerformanceRatingTable))
df <- plyr::rename(df, c("Var1"= "PerformanceRating", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(PerformanceRating, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
PerformanceRatingBP <- g + xlab("Performance Rating") + ylab("Percent") + ggtitle("Attrition vs. Performance Rating")
PerformanceRatingBP

```

### Attrition vs. Relationship Satisfaction
#### Relationship Satisfaction does not seem to correlate with attrition.

```{r Relationship Satisfaction, echo=FALSE}
# Generate property summary for variable
RelationshipSatisfactionTable <- table(cleanAttDf$RelationshipSatisfaction, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(RelationshipSatisfactionTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "RelationshipSatisfaction", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(RelationshipSatisfactionTable))
df <- plyr::rename(df, c("Var1"= "RelationshipSatisfaction", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(RelationshipSatisfaction, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
RelationshipSatisfactionBP <- g + xlab("Relationship Satisfaction") + ylab("Percent") + ggtitle("Attrition vs. Relationship Satisfaction")
RelationshipSatisfactionBP

```

### Attrition vs. Stock Option Level
#### Stock Option Level probably reflects status and age, indirectly correlated with attrition.

```{r Stock Option Level, echo=FALSE}
# Generate property summary for variable
StockOptionLevelTable <- table(cleanAttDf$StockOptionLevel, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(StockOptionLevelTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "StockOptionLevel", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(StockOptionLevelTable))
df <- plyr::rename(df, c("Var1"= "StockOptionLevel", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(StockOptionLevel, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
StockOptionLevelBP <- g + xlab("Stock Option Level") + ylab("Percent") + ggtitle("Attrition vs. Stock Option Level")
StockOptionLevelBP

```

### Attrition vs. Total Working Years
#### Attrition decreases with Total Working Years and this also may be related to age.

```{r Total Working Years, echo=FALSE}
## Subset Total Working Years by 5, categorical vector
# 0-4,5-9,10-14,15-19,20-24,25-29,30-34,35-40
cleanAttDf$TotalWorkingYearsby5 <- findInterval(cleanAttDf$TotalWorkingYears, c(0, 5, 10, 20, 25, 30, 35, 40))
cleanAttDf$TotalWorkingYearsby5 <- as.factor(cleanAttDf$TotalWorkingYearsby5)
cleanAttDf$TotalWorkingYearsby5 <- mapvalues(cleanAttDf$TotalWorkingYearsby5, from=c(1, 2, 3, 4, 5, 6, 7, 8), to=c('0-4', '5-9', '10-14', '15-19', '20-24', '25-29','30-34','35-40'))

# Generate property summary for variable
TotalWorkingYearsby5Table <- table(cleanAttDf$TotalWorkingYearsby5, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(TotalWorkingYearsby5Table))
PTdf <- plyr::rename(PTdf, c("Var1"= "TotalWorkingYearsby5", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(TotalWorkingYearsby5Table))
df <- plyr::rename(df, c("Var1"= "TotalWorkingYearsby5", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = TotalWorkingYearsby5, y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
TotalWorkingYearsby5BP <- g + xlab("Total Working Years") + ylab("Percent") + ggtitle("Attrition vs. Total Working Years")
TotalWorkingYearsby5BP

```

### Attrition vs. Training Times Last Year
#### Training Times Last Year does not seem to correlate with attrition.

```{r Training Times Last Year, echo=FALSE}
# Generate property summary for variable
TrainingTimesLastYearTable <- table(cleanAttDf$TrainingTimesLastYear, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(TrainingTimesLastYearTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "TrainingTimesLastYear", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(TrainingTimesLastYearTable))
df <- plyr::rename(df, c("Var1"= "TrainingTimesLastYear", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = (TrainingTimesLastYear), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
TrainingTimesLastYearBP <- g + xlab("Training Times Last Year") + ylab("Percent") + ggtitle("Attrition vs. Training Times Last Year")
TrainingTimesLastYearBP

```

### Attrition vs. Work Life Balance
#### Poor Work Life Balance seem to have some influence on increased attrition.

```{r Work Life Balance, echo=FALSE}
# Generate property summary for variable
WorkLifeBalanceTable <- table(cleanAttDf$WorkLifeBalance, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(WorkLifeBalanceTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "WorkLifeBalance", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(WorkLifeBalanceTable))
df <- plyr::rename(df, c("Var1"= "WorkLifeBalance", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = fct_reorder(WorkLifeBalance, Freq), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
WorkLifeBalanceBP <- g + xlab("Work Life Balance") + ylab("Percent") + ggtitle("Attrition vs. Work Life Balance")
WorkLifeBalanceBP

```

### Attrition vs. Years At Company
#### Attrition decreases with Years At Company and this too may be related to age.

```{r Years At Company, echo=FALSE}
## Subset Total Working Years by 5, categorical vector
cleanAttDf$YearsAtCompanyby5 <- findInterval(cleanAttDf$YearsAtCompany, c(0, 5, 10, 15, 20, 25, 30, 35))
cleanAttDf$YearsAtCompanyby5 <- mapvalues(cleanAttDf$YearsAtCompanyby5, from=c(1, 2, 3, 4, 5, 6, 7, 8), to=c('0-04', '05-09', '10-14', '15-19', '20-24', '25-29', '30-34', '35-40'))
cleanAttDf$YearsAtCompanyby5 <- as.factor(cleanAttDf$YearsAtCompanyby5)

# Generate property summary for variable
YearsAtCompanyby5Table <- table(cleanAttDf$YearsAtCompanyby5, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(YearsAtCompanyby5Table))
PTdf <- plyr::rename(PTdf, c("Var1"= "YearsAtCompanyby5", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(YearsAtCompanyby5Table))
df <- plyr::rename(df, c("Var1"= "YearsAtCompanyby5", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = YearsAtCompanyby5, y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
YearsAtCompanyby5BP <- g + xlab("Years At Company") + ylab("Percent") + ggtitle("Attrition vs. Years At Company")
YearsAtCompanyby5BP

```

### Attrition vs. Years In Current Role
#### Attrition decreases with Years In Current Role and this could be related to age.

```{r Years In Current Role, echo=FALSE}
# Generate property summary for variable
YearsInCurrentRoleTable <- table(cleanAttDf$YearsInCurrentRole, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(YearsInCurrentRoleTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "YearsInCurrentRole", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(YearsInCurrentRoleTable))
df <- plyr::rename(df, c("Var1"= "YearsInCurrentRole", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = (YearsInCurrentRole), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
YearsInCurrentRoleBP <- g + xlab("Years In Current Role") + ylab("Percent") + ggtitle("Attrition vs. Years In Current Role")
YearsInCurrentRoleBP

```

### Attrition vs. Years Since Last Promotion
#### Years Since Last Promotion seems to have little influence on attrition.

```{r Years Since Last Promotion, echo=FALSE}
# Generate property summary for variable
YearsSinceLastPromotionTable <- table(cleanAttDf$YearsSinceLastPromotion, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(YearsSinceLastPromotionTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "YearsSinceLastPromotion", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(YearsSinceLastPromotionTable))
df <- plyr::rename(df, c("Var1"= "YearsSinceLastPromotion", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = (YearsSinceLastPromotion), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
YearsSinceLastPromotionBP <- g + xlab("Years Since Last Promotion") + ylab("Percent") + ggtitle("Attrition vs. Years Since Last Promotion")
YearsSinceLastPromotionBP

```
### Attrition vs. Years With Current Manager
#### Years With Current Manager seems to have little influence on attrition beyond the first year.

```{r Years With Current Manager, echo=FALSE}
# Generate property summary for variable
YearsWithCurrManagerTable <- table(cleanAttDf$YearsWithCurrManager, cleanAttDf$Attrition)
PTdf <- as.data.frame(prop.table(YearsWithCurrManagerTable))
PTdf <- plyr::rename(PTdf, c("Var1"= "YearsWithCurrManager", "Var2"= "Attrition"))
df <- as.data.frame(prop.table(YearsWithCurrManagerTable))
df <- plyr::rename(df, c("Var1"= "YearsWithCurrManager", "Var2"= "Attrition"))

kable(df, "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
# Plot variable vs Attrition Percent
g <- ggplot(PTdf, aes(x = (YearsWithCurrManager), y = (100*Freq), fill=Attrition)) + geom_bar(stat = "identity")
YearsWithCurrManagerBP <- g + xlab("Years With Current Manager") + ylab("Percent") + ggtitle("Attrition vs. Years With Current Manager")
YearsWithCurrManagerBP

```
## Model Preparation

```{r Downsample}
# Downsample the 'Yes' Attrition
set.seed(7)
cleanAttDfDs <- downSample(x=cleanAttDf, y=cleanAttDf$Attrition)

cleanAttDfDs_cat <- cleanAttDfDs[,names(base::Filter(is.factor,cleanAttDfDs))]
cleanAttDfDs_con <- cleanAttDfDs[,c("Attrition",names(base::Filter(is.integer,cleanAttDfDs)))]

cleanAttDfDs_conCorr <- cor(cleanAttDfDs_con[,2:15])

# Correlation heatmap
ggplot(data=melt(cleanAttDfDs_conCorr), aes(x=Var1,y=Var2,fill=value)) + geom_tile(color="white") +
  theme(axis.text.x = element_text(angle=90, hjust = 1)) +
  scale_fill_gradient2(low="purple",high="Red",mid="#fcfcfc" ,
                       midpoint = 0, limit=c(-1,1), space = "Lab",
                       name="Pearson \nCorrelation") +
  ggtitle("Correlation Matrix for Continuous Variables") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )

#splom(cleanAttDfDs_con[,2:15], groups=cleanAttDfDs_con$Attrition, data=cleanAttDfDs_con, panel=panel.superpose)

#pairs(cleanAttDfDs_con[,2:15],col=cleanAttDfDs_con$Attrition)

#histograms for categorical data using cleaned original files

barPlot1 <- ggplot(data = melt(orig_dfCat[,1:7], id=c("Attrition")), mapping= aes(x=value, fill=Attrition)) + geom_histogram(stat="count")
barPlot1 <- barPlot1 + facet_grid(Attrition~variable, scales = "free")+ theme(axis.text.x = element_text(angle=90, hjust = 1)) + xlab("Categories") 

barPlot2 <- ggplot(data = melt(orig_dfCat[,c(1,8:13)], id=c("Attrition")), mapping= aes(x=value, fill=Attrition) ) + geom_histogram(stat="count")  
barPlot2 <- barPlot2 + facet_grid(Attrition~variable, scales = "free")+ theme(axis.text.x = element_text(angle=90, hjust = 1)) + xlab("Categories") 

barPlot3 <- ggplot(data = melt(orig_dfCat[,c(1,14:17)], id=c("Attrition")), mapping= aes(x=value, fill=Attrition) ) +geom_histogram(stat="count")  
barPlot3 <- barPlot3 + facet_grid(Attrition~variable, scales = "free")+ theme(axis.text.x = element_text(angle=90, hjust = 1)) + xlab("Categories")

barPlot1
barPlot2
barPlot3

```
### Model

```{r RForest}

#use the clean data set that is down sampled.
cleanAttDfDs
ncol(cleanAttDfDs)

clf_1 <- randomForest(cleanAttDfDs[,c(1,3:31)],cleanAttDfDs[,2])
impClf_1<- clf_1$importance

#creates attractive random forest graph
clfdf<-cbind(rownames(impClf_1),impClf_1)
colnames(clfdf)<-c("variable","score")
clfdf<-as.data.frame(clfdf,row.names = F)
clfdf$score<-as.numeric(levels(clfdf$score))[clfdf$score]
clfdf <- transform(clfdf, variable=reorder(clfdf$variable, -clfdf$score) ) 

colourCount = length(unique(clfdf$variable))
getPalette = colorRampPalette(brewer.pal(11, "RdYlGn"))

ggplot(data=clfdf, aes(x=reorder(clfdf$variable,clfdf$score),y=clfdf$score,fill=clfdf$variable)) + geom_bar(stat="identity") + coord_flip() +
  labs(title="Random Forest model predictive factors",x="Predictor",y="Score") + 
  theme(plot.title = element_text(hjust = 0.3), legend.position ="none") + scale_fill_manual(values = getPalette(colourCount))



#Test ROC Random Forest
num_obs=nrow(cleanAttDfDs)

train_idx <- sample(c(1:num_obs),size=num_obs*0.7,replace = FALSE)

clf_2 = randomForest(cleanAttDfDs[train_idx,c(1,3:31)],cleanAttDfDs[train_idx,2])
predicAtt <- predict(clf_2,cleanAttDfDs[-train_idx,c(1,3:31)], type="prob" )
plot(roc(cleanAttDfDs[-train_idx,2], as.numeric(predicAtt[,1])))

auc.clf2 <- roc(cleanAttDfDs[-train_idx,2], as.numeric(predicAtt[,1]),auc=TRUE)
auc.clf2$auc

nloops <- 100
cv.aucs <-c()
cv.top3pred <- c()
for (i in 1:nloops){
  train_idx <- sample(c(1:num_obs),size=num_obs*0.7,replace = FALSE)
  clf_3 = randomForest(cleanAttDfDs[train_idx,c(1,3:31)],cleanAttDfDs[train_idx,2])
  predicAtt <- predict(clf_3,cleanAttDfDs[-train_idx,c(1,3:31)], type="prob" )
  auc.clf3 <- roc(cleanAttDfDs[-train_idx,2], as.numeric(predicAtt[,1]),auc=TRUE)
  #plot(roc(cleanAttDfDs[-train_idx,2], as.numeric(predicAtt[,1])))
  cv.aucs[i]<-as.numeric(auc.clf3$auc)
  Pred_Results <- clf_3$importance
  sortedResults <- as.matrix(Pred_Results[order(Pred_Results[,1], decreasing = "TRUE"),])
  Top3Pred <- rownames(sortedResults)
  cv.top3pred<-rbind(cv.top3pred,Top3Pred[1:3])
}

ggplot(melt(cv.top3pred), mapping = aes(x=value,fill="count")) + geom_histogram(stat = "count") + 
  theme(axis.text.x = element_text(angle=90, hjust = 1),legend.position = "none") + 
  labs(title="Random Forest model top 3 predictor occurences",x="Predictor",y="Count") +
  scale_fill_manual(values="#71945A")


qplot(cv.aucs,geom="histogram",binwidth=.008,main = "Histogram for random forest model", fill=I("Blue"), col=I("Gray"), alpha=I(.8)) + geom_density(col=2)

cv.aucs
summary(cv.aucs)

# Result plot

sortedResultsHist <- as.data.frame(sortedResults)
sortedResultsHist <- rename(sortedResultsHist, "Importance" = V1)
sortedResultsHist["Factor"] <- rownames(sortedResultsHist)
sortedResultsTop3 <- dplyr::slice(sortedResultsHist, 1:3)

# ggplot Top 3 Factors 
# Barplot
ggplot(sortedResultsTop3, aes(x=Factor, fill=(sortedResultsTop3$Factor), y=Importance)) + 
  labs(title="Top 3 predictor results",x="Predictor",y="Importance") + geom_bar(stat = "identity") + 
  theme(plot.title = element_text(hjust = 0.4), legend.position ="none") + scale_fill_manual(values=c("#5F9EA0","#4682B4","#00BFFF"))




```


## Conclusion

