---
title: "6306 Case Study 2 - "
author: "Andrew Walch, Christopher Morgan, Luke Pierce, Eduardo Cantu Medellin"
date: "April 7, 2018"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}



knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Exploratory Data Analysis


A summary done on the data main data yields that some of the fields may not be needed:
  EmployeeCount: This shows 1 for all of them
  EmployeeNumber: This has to do with a random number
  Over18: All of them are over 18
  StandardHours: All records show 80 hours a week.





```{r EDA}

library(pROC)
library(ROCR)
library(randomForest)
library(ggplot2)
library(reshape2)
library(caret)
# Add libraries
library(dplyr)
library(RColorBrewer)
#library(lattice)
#library(sm)
# Load Data
df <- read.csv("CaseStudy2data.csv")
dfCat <- df[,names(base::Filter(is.factor,df))]
dfCon <- df[,c("Attrition",names(base::Filter(is.integer,df)))]
head(dfCat)
summary(df)


```



The data shows an inbalance in the number of people that leave the company vs people that stayed in the job. There are 237 samples with attrition 'Yes' and 1233 with attrition 'No'. The data will be down sampled for the Attrition "Yes" to match the level of Attrition "Yes"

Also, the data was scanned for 'NA' values. This scanned did not show any 'NA' values.

LEP: Added branch for categorical variables for histogram plots. The downSample function removes 9 categorical variables (see cleanAttDfDs before and after downSample).


```{r NA_Scan}
# NA per Field
NACount <- sapply(df,function(x) sum(is.na(x)))
NACount <- as.data.frame(NACount)
colnames(NACount)=c("CountOfNAs")
NACount


#remove reduntant fields

#remove reduntant fields
# 9 - Employee Count
# 10 - Employee Number
# 22 - Over18
# 27 - Standard Hours
cleanAttDf <- df[,c(-9,-10,-22,-27)]
colnames(cleanAttDf)

# change variables as factors
cleanAttDf$JobLevel <- as.factor(cleanAttDf$JobLevel)
cleanAttDf$JobInvolvement <- as.factor(cleanAttDf$JobInvolvement)
cleanAttDf$EnvironmentSatisfaction <- as.factor(cleanAttDf$EnvironmentSatisfaction)
cleanAttDf$Education <- as.factor(cleanAttDf$Education)
cleanAttDf$JobSatisfaction <- as.factor(cleanAttDf$JobSatisfaction)
cleanAttDf$RelationshipSatisfaction <- as.factor(cleanAttDf$RelationshipSatisfaction)
cleanAttDf$StockOptionLevel <- as.factor(cleanAttDf$StockOptionLevel)
cleanAttDf$PerformanceRating <- as.factor(cleanAttDf$PerformanceRating)
cleanAttDf$WorkLifeBalance <- as.factor(cleanAttDf$WorkLifeBalance)

##### Branch for categorical histograms
orig_cleanAttDf <- cleanAttDf
orig_dfCat <- orig_cleanAttDf[,names(base::Filter(is.factor,orig_cleanAttDf))]

# Downsample the 'Yes' Attrition
set.seed(7)
cleanAttDfDs <- downSample(x=cleanAttDf, y=cleanAttDf$Attrition)

cleanAttDfDs_cat <- cleanAttDfDs[,names(base::Filter(is.factor,cleanAttDfDs))]
cleanAttDfDs_con <- cleanAttDfDs[,c("Attrition",names(base::Filter(is.integer,cleanAttDfDs)))]

cleanAttDfDs_conCorr <- cor(cleanAttDfDs_con[,2:15])

# Correlation heatmap
ggplot(data=melt(cleanAttDfDs_conCorr), aes(x=Var1,y=Var2,fill=value)) + geom_tile(color="white") +
  theme(axis.text.x = element_text(angle=90, hjust = 1)) +
  scale_fill_gradient2(low="purple",high="Red",mid="#fcfcfc" ,
                       midpoint = 0, limit=c(-1,1), space = "Lab",
                       name="Pearson \nCorrelation") +
  ggtitle("Correlation Matrix for Continuous Variables") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )

#splom(cleanAttDfDs_con[,2:15], groups=cleanAttDfDs_con$Attrition, data=cleanAttDfDs_con, panel=panel.superpose)

#pairs(cleanAttDfDs_con[,2:15],col=cleanAttDfDs_con$Attrition)

#histograms for categorical data using cleaned original files

barPlot1 <- ggplot(data = melt(orig_dfCat[,1:7], id=c("Attrition")), mapping= aes(x=value, fill=Attrition)) + geom_histogram(stat="count")
barPlot1 <- barPlot1 + facet_grid(Attrition~variable, scales = "free")+ theme(axis.text.x = element_text(angle=90, hjust = 1)) + xlab("Categories") 

barPlot2 <- ggplot(data = melt(orig_dfCat[,c(1,8:13)], id=c("Attrition")), mapping= aes(x=value, fill=Attrition) ) + geom_histogram(stat="count")  
barPlot2 <- barPlot2 + facet_grid(Attrition~variable, scales = "free")+ theme(axis.text.x = element_text(angle=90, hjust = 1)) + xlab("Categories") 

barPlot3 <- ggplot(data = melt(orig_dfCat[,c(1,14:17)], id=c("Attrition")), mapping= aes(x=value, fill=Attrition) ) +geom_histogram(stat="count")  
barPlot3 <- barPlot3 + facet_grid(Attrition~variable, scales = "free")+ theme(axis.text.x = element_text(angle=90, hjust = 1)) + xlab("Categories")

barPlot1
barPlot2
barPlot3

```
### Model

```{r RForest}

#use the clean data set that is down sampled.
cleanAttDfDs
ncol(cleanAttDfDs)

clf_1 <- randomForest(cleanAttDfDs[,c(1,3:31)],cleanAttDfDs[,2])
impClf_1<- clf_1$importance

#creates attractive random forest graph
clfdf<-cbind(rownames(impClf_1),impClf_1)
colnames(clfdf)<-c("variable","score")
clfdf<-as.data.frame(clfdf,row.names = F)
clfdf$score<-as.numeric(levels(clfdf$score))[clfdf$score]
clfdf <- transform(clfdf, variable=reorder(clfdf$variable, -clfdf$score) ) 

colourCount = length(unique(clfdf$variable))
getPalette = colorRampPalette(brewer.pal(11, "RdYlGn"))

ggplot(data=clfdf, aes(x=reorder(clfdf$variable,clfdf$score),y=clfdf$score,fill=clfdf$variable)) + geom_bar(stat="identity") + coord_flip() +
  labs(title="Random Forest model predictive factors",x="Predictor",y="Score") + 
  theme(plot.title = element_text(hjust = 0.3), legend.position ="none") + scale_fill_manual(values = getPalette(colourCount))



#Test ROC Random Forest
num_obs=nrow(cleanAttDfDs)

train_idx <- sample(c(1:num_obs),size=num_obs*0.7,replace = FALSE)

clf_2 = randomForest(cleanAttDfDs[train_idx,c(1,3:31)],cleanAttDfDs[train_idx,2])
predicAtt <- predict(clf_2,cleanAttDfDs[-train_idx,c(1,3:31)], type="prob" )
plot(roc(cleanAttDfDs[-train_idx,2], as.numeric(predicAtt[,1])))

auc.clf2 <- roc(cleanAttDfDs[-train_idx,2], as.numeric(predicAtt[,1]),auc=TRUE)
auc.clf2$auc

nloops <- 100
cv.aucs <-c()
cv.top3pred <- c()
for (i in 1:nloops){
  train_idx <- sample(c(1:num_obs),size=num_obs*0.7,replace = FALSE)
  clf_3 = randomForest(cleanAttDfDs[train_idx,c(1,3:31)],cleanAttDfDs[train_idx,2])
  predicAtt <- predict(clf_3,cleanAttDfDs[-train_idx,c(1,3:31)], type="prob" )
  auc.clf3 <- roc(cleanAttDfDs[-train_idx,2], as.numeric(predicAtt[,1]),auc=TRUE)
  #plot(roc(cleanAttDfDs[-train_idx,2], as.numeric(predicAtt[,1])))
  cv.aucs[i]<-as.numeric(auc.clf3$auc)
  Pred_Results <- clf_3$importance
  sortedResults <- as.matrix(Pred_Results[order(Pred_Results[,1], decreasing = "TRUE"),])
  Top3Pred <- rownames(sortedResults)
  cv.top3pred<-rbind(cv.top3pred,Top3Pred[1:3])
}

ggplot(melt(cv.top3pred), mapping = aes(x=value,fill="count")) + geom_histogram(stat = "count") + 
  theme(axis.text.x = element_text(angle=90, hjust = 1),legend.position = "none") + 
  labs(title="Random Forest model top 3 predictor occurences",x="Predictor",y="Count") +
  scale_fill_manual(values="#71945A")


qplot(cv.aucs,geom="histogram",binwidth=.008,main = "Histogram for random forest model", fill=I("Blue"), col=I("Gray"), alpha=I(.8)) + geom_density(col=2)

cv.aucs
summary(cv.aucs)

# Result plot

sortedResultsHist <- as.data.frame(sortedResults)
sortedResultsHist <- rename(sortedResultsHist, "Importance" = V1)
sortedResultsHist["Factor"] <- rownames(sortedResultsHist)
sortedResultsTop3 <- dplyr::slice(sortedResultsHist, 1:3)

# ggplot Top 3 Factors 
# Barplot
ggplot(sortedResultsTop3, aes(x=Factor, fill=(sortedResultsTop3$Factor), y=Importance)) + 
  labs(title="Top 3 predictor results",x="Predictor",y="Importance") + geom_bar(stat = "identity") + 
  theme(plot.title = element_text(hjust = 0.4), legend.position ="none") + scale_fill_manual(values=c("#5F9EA0","#4682B4","#00BFFF"))




```


## Conclusion

