---
title: "6306 Case Study 2 - "
author: "Andrew Walch, Christopher Morgan, Luke Pierce, Eduardo Cantu Medellin"
date: "April 7, 2018"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Exploratory Data Analysis


A summary done on the data main data yields that some of the fields may not be needed:
  EmployeeCount: This shows 1 for all of them
  EmployeeNumber: This has to do with a random number
  Over18: All of them are over 18
  StandardHours: All records show 80 hours a week.





```{r EDA}
library(pROC)
library(randomForest)
library(ggplot2)
library(reshape2)
# Load Data
df <- read.csv("CaseStudy2data.csv")


dfCat <- df[,names(base::Filter(is.factor,df))]

dfCon <- df[,c("Attrition",names(base::Filter(is.integer,df)))]
head(dfCat)


summary(df)
ggplot(data = melt(dfCat, id=c("Attrition")), mapping = aes(x = value)) + geom_histogram(stat="count") + facet_grid(Attrition~variable, scales = 'free') + ggtitle("Histograms for Categorical Variables by Attrition") + theme(axis.text.x = element_text(angle=90, hjust = 1)) + xlab("Categories") 

ggplot(data = melt(dfCon[,1:6], id=c("Attrition")), mapping = aes(x = value)) + geom_histogram(bin=15) + facet_grid(Attrition~variable, scales = 'free')+ ggtitle("Histograms for Continuous Variables by Attrition")

ggplot(data = melt(dfCon[,c(1,7:11)], id=c("Attrition")), mapping = aes(x = value)) + geom_histogram(bin=15) + facet_grid(Attrition~variable, scales = 'free')+ ggtitle("Histograms for Continuous Variables by Attrition")

ggplot(data = melt(dfCon[,c(1,12:16)], id=c("Attrition")), mapping = aes(x = value)) + geom_histogram(bin=15) + facet_grid(Attrition~variable, scales = 'free')+ ggtitle("Histograms for Continuous Variables by Attrition")

ggplot(data = melt(dfCon[,c(1,17:21)], id=c("Attrition")), mapping = aes(x = value)) + geom_histogram(bin=15) + facet_grid(Attrition~variable, scales = 'free')+ ggtitle("Histograms for Continuous Variables by Attrition")

ggplot(data = melt(dfCon[,c(1,22:27)], id=c("Attrition")), mapping = aes(x = value)) + geom_histogram(bin=15) + facet_grid(Attrition~variable, scales = 'free')+ ggtitle("Histograms for Continuous Variables by Attrition")


```

Also, the data was scanned for 'NA' values. This scanned did not show any 'NA' values.
```{r NA_Scan}
# NA per Field
NACount <- sapply(df,function(x) sum(is.na(x)))
NACount <- as.data.frame(NACount)
colnames(NACount)=c("CountOfNAs")
NACount


```
### Charts



## Conclusion

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
